Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-27 algorithm.tex
AA: aggiustata tutta la prima parte

\item The total innovation and average score for the entire era is calculated:
	\[
	is_{TOT,i}=\sum_{R\in\mathcal{R}_{i}}is\left(R\right)
	\]

\item The configuration space partition $\mathcal{R}_{i}$ is divided in three subsets, that we will comment later \comment{e invece sarebbe bene commentarli ora, anche}:

	\begin{enumerate}
	\item high innovation regions 
	\[
	\mathcal{R}_{i,h}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)>\alpha\cdot is_{av,i-1}\right\} 
	\]
	 where $\alpha$ is a constant defined by the designer (for example
	$\alpha=1.2$). 
	\item low innovation regions 
	\[
	\mathcal{R}_{i,l}=\left\{ \left.R\in\mathcal{R}_{i}\right|0<is\left(R\right)\le\alpha\cdot is_{av,i-1}\right\} 
	\]
	\item no innovation regions 
	\[
	\mathcal{R}_{i,n}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)=0\right\} 
	\]
	\end{enumerate}

\item The goal of this step is preparing the regions that will compose the configuration space partition of the next era, i.e. $\mathscr{R}_{i+1}$. The following operations are performed:
	\begin{enumerate}
		\item Each high innovation region $R\in\mathscr{R}_{i,h}$ is split obtaining $\psi(R)$, as stated in def \defR{splitting-a-region}. We denote with $\psi\left(\mathscr{R}_{i,h}\right)$ the set of regions obtained as explained.
		\item Low innovation regions are left unchanged.
		\item No innovation regions are merged, two by two. To do so, couples of no innovation contiguous regions $(R_1,R_2)$ are selected. We denote the set of these couples as $\mathscr{M}_i$. This set is constructed such that couples are independent, i.e. there cannot be two of them sharing a region. In other words, taking whatever $(R_1,R_2),(R_3,R_4) \in \mathscr{M}_i$, $R_i \neq R_j, \forall i,j\in\lbrace1,2,3,4\rbrace, i\neq$. 
We merge each couple of regions, i.e. $\forall (R_1,R_2)\in\mathscr{M}_i$ we calculate $\mu(R_1,R_2)$, according to \defR{merging-regions}.
We stress that the regions that belong to each selected couple must be contiguous, otherwise it is impossible to merge them. We denote with $\mu(\mathscr{M}_i)$ the set of the merged regions, i.e. $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$
	\end{enumerate}

\item The set of regions $\mathcal{R}_{i+1}$ for the following era is computed as:
	\[
	\mathcal{R}_{i+1}=
	\]
	\[
	=\psi\left(\mathscr{R}_{i,h}\right) \cup \mathcal{R}_{i,l}\cup
	\]
	\[
	\cup\left(\mathcal{R}_{i,n}\setminus\bigcup_{\left(R_{1},R_{2}\right)\in\mathcal{C}}\left\{ R_{1},R_{2}\right\} \right)\cup \mu(\mathscr{M}_i)
	\]
\end{enumerate}

\comment{commentare questa formula}

\comment{SONO ARRIVATO A LEGGERE QUA}

\subsection{Termination Condition}
The algorithm terminates at step $k$ if
\[
is_{TOT,k},is_{TOT,k-1},\dots,is_{TOT,k-\beta}\le\gamma\cdot is_{TOT,k-\left(\beta-1\right)}
\]
 where $\beta$ and $\gamma$ are selected by the experimenter (see
remark \ref{rem:termination}).
\begin{remark}
\label{pers02.rem:smaller_regions}The rule \ref{pers02.enu:K} of
the iteration rules says that the smaller is a region, the more ``crowded''
it is. It means that configurations in smaller regions are evaluated
with more ``attention'' (see figure \ref{pers02.fig:small_and_big}).
\end{remark}

\begin{remark}
The rule \ref{pers02.enu:novelty_score_of_a_configuration} says that
a configuration whose objectives are more ``separated'' by the previous
Pareto front, has to be taken in greater account than others. A configuration
that is very near, in the objective space, to the previous Pareto
front doesn't add a true innovation. Therefore, the innovation score
of a configuration is a measure of how much interest we have in considering
it (see figure \ref{pers02.fig:Novelty-score-of-a-config}).

\begin{figure}[h]
\includegraphics[width=0.5\columnwidth]{img/novelty_score}

\caption{\label{pers02.fig:Novelty-score-of-a-config}innovation score of a
configuration}

\end{figure}

\comment{Ci vorrebbe una rappresentazione schematica dell'algoritmo usando l'ambiente latex Algorithm o non mi ricordo come si chiama}

\end{remark}

\begin{remark}
The innovation score of a region is a measure of how much different
are the points that it adds to Pareto Front compared to the previous
Pareto front. ``High innovation regions'' are the ones in which
more configurations are worth evaluating, because they add many points
on not yet well explored Pareto front areas. Therefore, these regions
are split in the next era (see remark \ref{pers02.rem:smaller_regions})
and more ``crowded'' by simulations. ``No innovation regions''
are the ones that have proved to add few points on Pareto fronts,
probably very near to previous Pareto front points and a Pareto front
point near to previous points is not so much interesting. Therefore
it's not worth evaluating many of their configurations and it is useful
to merge these regions, in order to less densely evaluate them.

%ULTERIORI GIUSTIFICAZIONI SULL'IMPORTANZA DI CONSIDERARE LA DIVERSITA'
%DELLE SOLUZIONI SI TROVA IN ext12.pag5 E ALTRI.

\end{remark}

\begin{remark}
\label{rem:termination}The algorithm terminates when in the last
$\beta$ iterations not a great ``amount of innovation'' has been
discovered. Therefore it is very probably that carrying on iterating
will not considerably change the Pareto front already formed. It's
not recommended to chose $\beta=1$, because it is possible that the
Pareto fronts $\mathbf{o}\left(\mathscr{P}_{k-1}\right)$ and $\mathbf{o}\left(\mathscr{P}_{k}\right)$
are not very different but next Pareto fronts could be. In other words,
it's unsafe to terminate as soon as the Pareto front doesn't change
enough between two consecutive eras, thus a strategy taking into
account the history of the Pareto fronts has been adopted in the
proposed approach.
\end{remark}
%!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
%FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

\begin{remark}
Merging and splitting the regions are very dynamic processes. A portion
of the parameter space can be populated by a large number of small
regions in some eras, but can be covered by a small number of large
regions in next eras. This can be explained saying that, after a number
of eras of intense exploration of a portion of parameter space, this
exploration has turned to be enough thorough and so it was time for
other portions to be minutely scanned.
\end{remark}

>>>>

\item The total innovation and average score for the entire era is calculated:
	\begin{align} \begin{array}{l}
	is_{TOT,i}=\sum_{R\in\mathcal{R}_{i}}is\left(R\right) \\
	is_{av,i}=is_{TOT,i} / |\mathcal{R}_i|
	\end{array} \end{align}

\item The configuration space partition $\mathcal{R}_{i}$ is divided in three subsets, that we will comment later:

	\begin{enumerate}
	\item \emph{high innovation regions}, whose innovation score exceeds the average score of the previous era by a factor of $\alpha$, i.e.
	\[
	\mathcal{R}_{i,h}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)>\alpha\cdot is_{av,i-1}\right\} 
	\]
	 where $\alpha$ is a constant defined by the designer (for example
	$\alpha=1.2$). 
	\item \emph{low innovation regions}, whose score is positive but modest, i.e.
	\[
	\mathcal{R}_{i,l}=\left\{ \left.R\in\mathcal{R}_{i}\right|0<is\left(R\right)\le\alpha\cdot is_{av,i-1}\right\} 
	\]
	\item \emph{no innovation regions}, whose innovation is null, i.e. 
	\[
	\mathcal{R}_{i,n}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)=0\right\} 
	\]
	\end{enumerate}

\item The goal of this step is preparing the regions that will compose the configuration space partition of the next era, i.e. $\mathscr{R}_{i+1}$. The following operations are performed:
	\begin{enumerate}
		\item Each high innovation region $R\in\mathscr{R}_{i,h}$ is split obtaining $\psi(R)$, as stated in def \defR{splitting-a-region}. We denote with $\psi\left(\mathscr{R}_{i,h}\right)$ the set of regions obtained as explained.
		\item Low innovation regions are left unchanged.
		\item No innovation regions are merged, two by two. To do so, pairs of no innovation contiguous regions $(R_1,R_2)$ are selected. We denote the set of these pairs as $\mathscr{M}_i$. This set is constructed such that pairs are independent, i.e. there cannot be two of them sharing a region. In other words, taking whatever $(R_1,R_2),(R_3,R_4) \in \mathscr{M}_i$, $R_i \neq R_j, \forall i,j\in\lbrace1,2,3,4\rbrace, i\neq$. 
We merge each pair of regions, i.e. $\forall (R_1,R_2)\in\mathscr{M}_i$ we calculate $\mu(R_1,R_2)$, according to \defR{merging-regions}.
We stress that the regions that belong to each selected pair must be contiguous, otherwise it is impossible to merge them. We denote with $\mu(\mathscr{M}_i)$ the set of the merged regions, i.e. $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$
	\end{enumerate}

\item The set of regions $\mathcal{R}_{i+1}$ for the following era is composed of the subregions resulting from splitting operations ($\psi\left(\mathscr{R}_{i,h}\right)$), the low innovation regions unchanged $\mathcal{R}_{i,l}$, the result of the merging operations ($\mu(\mathscr{M}_i))$ and the no innovation regions that was not possible to merge $\left(\mathcal{R}_{i,n}\setminus\bigcup_{\left(R_{1},R_{2}\right)\in\mathcal{C}}\left\{ R_{1},R_{2}\right\} \right)$. Formally:
	\[
	\mathcal{R}_{i+1}=
	\]
	\[
	=\psi\left(\mathscr{R}_{i,h}\right) \cup \mathcal{R}_{i,l}\cup
	\]
	\[
	\cup\left(\mathcal{R}_{i,n}\setminus\bigcup_{\left(R_{1},R_{2}\right)\in\mathcal{C}}\left\{ R_{1},R_{2}\right\} \right)\cup \mu(\mathscr{M}_i)
	\]
\end{enumerate}

\subsection{Termination Condition}
The algorithm terminates at step $k$ if
\[
is_{TOT,k},is_{TOT,k-1},\dots,is_{TOT,k-\beta}\le\gamma\cdot is_{TOT,k-\left(\beta-1\right)}
\]
 where $\beta$ and $\gamma$ are selected by the experimenter. 
The algorithm terminates when in the last
$\beta$ iterations not a great ``amount of innovation'' has been
discovered. Therefore it is very probably that carrying on iterating
will not considerably change the Pareto-front approximation already formed.

It is not recommended to chose $\beta=1$, because it is possible that the
$k-1$-th and $k$-th Pareto front approximations are not very different but next ones could be. In other words,
it is unsafe to terminate as soon as the Pareto front does not considerably change
between two consecutive eras, thus a strategy taking into
account the history of the Pareto fronts has been adopted in the
proposed approach.
%!!! IN INGLESE E' PIU' CORRETTO DIRE: ``IT IS POSSIBLE THAT THE PARETO
%FRONTS WERE NOT VERY DIFFERENT''? (CON ``WERE'' AL POSTO DI ``ARE'')

Observe that merging and splitting operation are dynamic processes. A portion
of the parameter space may be populated by a large number of small
regions in some eras, but can be covered by a small number of large
regions in next eras. This is a desirable property of PS, since, after a number
of eras of intense exploration of a portion of configuration space, this
exploration has turned to be enough thorough and so it may be time for
other portions to be minutely scanned.
\end{remark}
